\documentclass[12pt]{article}
\begin{document}
In the article, Brooks argues that there have not been any major breakthroughs in the software development field that solves the essential difficulties that follows software engineering.\\
Brooks identifies the essential difficulties in software engineering as complexity, conformity, changeability, invisibility.\\
Brooks view the complexity as an essential property of software hence descriptions of a software entity that abstract away its complexity often abstract away its essence. The complexity is not only an issue of technicality but also management problems stem from it.\\
As a software engineer Brooks states that since there is no divine being behind the evolution of software development techniques, we must rely on the conformity of prior endeavours, which only adds to the complexity of software as they have to conform to other interfaces.\\
Software does not have a physical form, therefore it is unvisualizeable, he claims. This is a problem because its impends the process of design and it hinders the communication among developers.\\ \\
Brooks lists few "solutions" that may act as a silver bullet in the software engineering industry. These include higher-level language advances, object-oriented programming, AI, expert systems, "automatic" programming, graphical programming, programming verification and workstations. He also has some suggestion on what could be promising attacks on the conceptual essence. These are buy versus build, requirement refinement and rapid prototyping, incremental development-grow, not build software and great designers.\\ \\
A point we can relate to, is the incremental development-grow, not build software part he discusses. The idea of the scale of the projects that grows alongside  our skills, gives the continuous sense of accomplishment, which benefits the group. As Brooks says the moral effects are startling, our groups enthusiasm loading up the first front end page was remarkable. Further the incremental development allows for the application to morph unevenly as members explores their talents with aspects of the android application.\\

\end{document}